\tealeaves: an R package for modeling leaf temperature using energy budgets

### 2019-01-24 ms/ms.Rnw
added figure to ms

@@

Evaporation rate (mol H$_2$O m$^{-2}$ s$^{-1}$) is the product of the total conductance to water vapour (Equation \ref{eq:g_tw}) and the water vapour gradient (Equation \ref{eq:d_wv}):

@@

R is a fully open source programming language for statistical computing that allows users to develop their own packages with new functions. \tealeaves~takes three sets of parameter inputs: leaf parameters, environmental parameters, and physical constants (see Table \ref{table:table_input}). The package provides reasonable defaults, but users can input new values to address their question, as I demonstrate in the next section. With one or more parameter sets, \tealeaves~uses the \code{uniroot} function in R base package \pkg{stats} to find the \tleaf~that balances the leaf energy budget (Equation \ref{eq:energy_budget}). It outputs the equilibrium \tleaf~and energy fluxes in a table for analysis and visualization.

@@

% limitations and future directions
Ultimately, the goal of \tealeaves~is to provide a platform for implementing very complex and fully customizable energy budget models. Such models may take too much computational time to be be useful for large-scale ecosystem models, but they can help understand a wider range of fascinating and poorly understood leaf anatomical and morphological features, as well as identify under what conditions simpler leaf temperature models are adequate. Currently, \tealeaves~has several limitations that I plan to address in future releases. It has rather simple models of infrared radiation and direct versus diffuse radiation. Ideally, it would be better if users could supply their own functions to calculate these parameters from the total irradiance. The model also assumes leaves are horizontal, whereas leaf orientation varies widely. Following previous authors, I modeled heat transfer as a mixed convection (Equations \ref{eq:nusselt} and \ref{eq:sherwood}, but this may not adequately describe real leaf behavior \citep{Roth-Nebelsick_2001}. \tealeaves~calculates equilibrium as opposed to transient behavior, which may takes several minutes to reach. Finally, the model assumes a single homogenous leaf temperature rather than using finite element modeling to calculate leaf temperature gradients across leaves of different shapes. These are important limitations of the current software which can be addressed in future work.

@@

@@

\label{table:table_output}
\end{table}
\end{document}

>>>>

\section*{Summary }

@@

Evaporation rate (mol H$_2$O m$^{-2}$ s$^{-1}$) is the product of the total conductance to water vapour (Eqn \ref{eq:g_tw}) and the water vapour gradient (Eqn \ref{eq:d_wv}):

@@

R is a fully open source programming language for statistical computing that allows users to develop their own packages with new functions. \tealeaves~takes three sets of parameter inputs: leaf parameters, environmental parameters, and physical constants (see Table \ref{table:table_input}). The package provides reasonable defaults, but users can input new values to address their question, as I demonstrate in the next section. With one or more parameter sets, \tealeaves~uses the \code{uniroot} function in R base package \pkg{stats} to find the \tleaf~that balances the leaf energy budget (Eqn \ref{eq:energy_budget}). It outputs the equilibrium \tleaf~and energy fluxes in a table for analysis and visualization.

@@

% limitations and future directions
Ultimately, the goal of \tealeaves~is to provide a platform for implementing very complex and fully customizable energy budget models. Such models may take too much computational time to be be useful for large-scale ecosystem models, but they can help understand a wider range of fascinating and poorly understood leaf anatomical and morphological features, as well as identify under what conditions simpler leaf temperature models are adequate. Currently, \tealeaves~has several limitations that I plan to address in future releases. It has rather simple models of infrared radiation and direct versus diffuse radiation. Ideally, it would be better if users could supply their own functions to calculate these parameters from the total irradiance. The model also assumes leaves are horizontal, whereas leaf orientation varies widely. Following previous authors, I modeled heat transfer as a mixed convection (Eqns \ref{eq:nusselt} and \ref{eq:sherwood}, but this may not adequately describe real leaf behavior \citep{Roth-Nebelsick_2001}. \tealeaves~calculates equilibrium as opposed to transient behavior, which may takes several minutes to reach. Finally, the model assumes a single homogenous leaf temperature rather than using finite element modeling to calculate leaf temperature gradients across leaves of different shapes. These are important limitations of the current software which can be addressed in future work.

@@

%--------------------------------------------------
% Figures
%--------------------------------------------------

\begin{figure}[ht]
\centerline{\includegraphics[width=\textwidth,height=\textwidth]{../figures/fig1.pdf}}
\end{figure}

\begin{figure} [t!]
\caption{Extended examples of \tealeaves. Code to generate these examples is provided in the Supporting Information. A. The temperature of smaller leaves is more closely coupled to air temperature. Each line represents a different leaf size (small, dashed line; medium, dotted line; large, solid line) and the leaf-to-air temperature differential ($y$-axis) over a gradient of air temperature ($x$-axis). B. Greater stomatal conductance cools leaves. Each line represents a different stomatal conductance (low, dashed line, 1 [$\mu$mol m$^{-2}$ s$^{-1}$ Pa$^{-1}$]; medium, dotted line, 3 [$\mu$mol m$^{-2}$ s$^{-1}$ Pa$^{-1}$]; high, solid line, 5 [$\mu$mol m$^{-2}$ s$^{-1}$ Pa$^{-1}$]) and the leaf-to-air temperature differential ($y$-axis) over a gradient of irradiance ($x$-axis). C. Forced convection dominates in small leaves; free convection dominates in very large leaves. Leaf size is indicated by line type as in Panel A. Horizontal lines indicate approximate shifts from forced convection ($Ar < 0.1$), mixed convection ($0.1 < Ar < 10$), and free convection ($Ar > 10$). Small leaves always experience forced convection, leading to lower leaf temperature compared to large leaves experiencing free convection. D. Amphistomatous leaves (Stomatal Ratio $\sim$ 0.5) evaporate more than hypo- or hyperstomatous leaves (Stomatal Ratio $\sim$ 0 or 1, respectively), especially under free convection (low wind speed, $u$).}
\label{fig:fig1}
\end{figure}

@@

\label{table:table_output}
\end{table}

%--------------------------------------------------
% Supporting Information
%--------------------------------------------------

\section*{Supporting Information}

R code for running extended examples. 
<<ExtendedExample1, echo = TRUE, eval = FALSE, results = hide>>=

# Extended example 1: 
# leaf size and leaf-to-air temperature differential

cs <- make_constants()

lp  <- make_leafpar(
  replace = list(
    leafsize = set_units(c(0.005, 0.1, 0.5), "m")
  )
)

ep <- make_enviropar(
  replace = list(
    S_sw = set_units(660, "W/m^2"),
    T_air = set_units(seq(278.15, 308.15, 5), "K")
  )
)

exe1 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
                unitless = TRUE, parallel = TRUE)

<<ExtendedExample2, echo = TRUE, eval = FALSE, results = hide>>=

# Extended example 2: 
# Solar radiation and leaf-to-air temperature differential

lp  <- make_leafpar(
  replace = list(
    g_sw = set_units(c(1, 3, 5), "umol/m^2/s/Pa")
  )
)

ep <- make_enviropar(
  replace = list(
    S_sw = set_units(seq(50, 950, 100), "W/m^2")
  )
)

exe2 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
                unitless = TRUE, parallel = TRUE)

<<ExtendedExample3, echo = TRUE, eval = FALSE, results = hide>>=

# Extended example 3: 
# Wind speed and leaf-to-air temperature differential

lp  <- make_leafpar(
  replace = list(
    leafsize = set_units(c(0.005, 0.1, 0.5), "m")
  )
)

ep <- make_enviropar(
  replace = list(
    wind = set_units(exp(seq(log(0.01), log(10), 
                             length.out = 1e2)), "m/s")
  )
)

exe3 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
                unitless = TRUE, parallel = TRUE)

<<ExtendedExample4, echo = TRUE, eval = FALSE, results = hide>>=

# Extended example 4: 
# Stomatal ratio and evaporation

lp  <- make_leafpar(
  replace = list(
    g_sw = set_units(c(0.4, 4), "umol/s/m^2/Pa"),
    logit_sr = set_units(seq(-10, 10, length.out = 1e2))
  )
)

ep <- make_enviropar(
  replace = list(
    RH = set_units(0.2),
    T_air = set_units(293.15, "K"),
    wind = set_units(c(0, 2), "m/s")
  )
)

exe4 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
                unitless = TRUE, parallel = TRUE)

\end{document}

### 2019-01-24 ms/ms.tex
added figure to ms

<<Import-objects, echo = FALSE, results = hide>>=

  setwd("..")
  source("r/header.R")

  # import2ms()
  setwd("ms")

@

@@

\label{table:table_output}
\end{table}
\end{document}

>>>>

@@

\label{table:table_output}
\end{table}

%--------------------------------------------------
% Supporting Information
%--------------------------------------------------

\section*{Supporting Information}

R code for running extended examples. 
\begin{Schunk}
\begin{Sinput}
> # Extended example 1: 
> # leaf size and leaf-to-air temperature differential
> 
> library(tealeaves)
> cs <- make_constants()
> lp  <- make_leafpar(
+   replace = list(
+     leafsize = set_units(c(0.005, 0.1, 0.5), "m")
+   )
+ )
> ep <- make_enviropar(
+   replace = list(
+     S_sw = set_units(660, "W/m^2"),
+     T_air = set_units(seq(278.15, 308.15, 5), "K")
+   )
+ )
> exe1 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
+                 unitless = TRUE, parallel = TRUE)
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> # Extended example 2: 
> # Solar radiation and leaf-to-air temperature differential
> 
> library(tealeaves)
> lp  <- make_leafpar(
+   replace = list(
+     g_sw = set_units(c(1, 3, 5), "umol/m^2/s/Pa")
+   )
+ )
> ep <- make_enviropar(
+   replace = list(
+     S_sw = set_units(seq(50, 950, 100), "W/m^2")
+   )
+ )
> exe2 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
+                 unitless = TRUE, parallel = TRUE)
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> # Extended example 3: 
> # Wind speed and leaf-to-air temperature differential
> 
> library(tealeaves)
> lp  <- make_leafpar(
+   replace = list(
+     leafsize = set_units(c(0.005, 0.1, 0.5), "m")
+   )
+ )
> ep <- make_enviropar(
+   replace = list(
+     wind = set_units(exp(seq(log(0.01), log(10), 
+                              length.out = 1e2)), "m/s")
+   )
+ )
> exe3 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
+                 unitless = TRUE, parallel = TRUE)
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> # Extended example 4: 
> # Stomatal ratio and evaporation
> 
> library(tealeaves)
> lp  <- make_leafpar(
+   replace = list(
+     g_sw = set_units(c(0.4, 4), "umol/s/m^2/Pa"),
+     logit_sr = set_units(seq(-10, 10, length.out = 1e2))
+   )
+ )
> ep <- make_enviropar(
+   replace = list(
+     RH = set_units(0.2),
+     T_air = set_units(293.15, "K"),
+     wind = set_units(c(0, 2), "m/s")
+   )
+ )
> exe4 <- tleaves(lp, ep, cs, progress = TRUE, quiet = TRUE, 
+                 unitless = TRUE, parallel = TRUE)
> 
\end{Sinput}
\end{Schunk}

\end{document}
